Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-04-25 paper/paper2_neo/paper.tex
Change paper and create multi-reconstructor

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into an interpretable format.
We wanted to represent the spectra as a list of spectral lines rather than a
list of fluxes. This was done using \texttt{blase}, which detects spectral 
lines as Voigt profiles and tunes the profiles to mimic the original 
PHOENIX spectrum with back propagation. Four parameters were optimized: the 
line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
learning rate of 0.05 over 100 epochs. In addition, we limited two custom 
parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
that determines the extent of the Voigt profile to evaluate, saving
computational resources by not evaluating small numbers. Prominence sets a
lower limit for the amplitude of detected lines, which also saves resources 
by disregarding shallow lines. In short, larger values for wing cut and 
smaller values for prominence both increase the accuracy of using
\texttt{blase}'s line detection and optimization at the expense of an
increased computational cost.

Once optimization was complete, the list of identified lines, present in the
state dictionary of \texttt{blase}'s model, was saved to a \texttt{.pt} file
for each PHOENIX subset grid point. The total disk space these files took up 
is 465 MB.

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
Earlier we mentioned that \texttt{blase} tuned the line centers of
detected lines. This means that from one PHOENIX spectrum to the next, the
same line could have a slightly different line center. Since the goal of this
study is to interpolate the properties of each line, we needed to identify the
presence of a particular line across the PHOENIX subset. We decided to do
this by using the line centers $\mu$ of the detected lines pre-optimization, 
while the post-optimization line centers are labeled $\mu'$. Now with 
each spectral line indexed by $\mu$, we had four parameters to interpolate:
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Mathematically, the Voigt profile
$\mathsf{V}_\mu$ in terms of our four parameters is given by:

\begin{gather}
    \mathsf{V}_\mu(\lambda) = e^{\ln(a)}\mathrm{Re}\left[w\left(\dfrac{\lambda - \mu' + i\gamma}{\sigma\sqrt{2}}\right)\right] \bigg/ \left(\sigma\sqrt{2\pi}\right)
\end{gather}
where $w$ is the Faddeeva function. Note that since we are dealing with 
the parameters of these Voigt profiles, even if the interpolation method is 
linear, the actual result of interpolation varies non-linearly.

@@

\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{images/figure2.png}
    \caption{Heatmap showing how $\ln(a)$ varies over the PHOENIX subset
    slice at solar metallicity. Notice the missing chunk in the top left
    of the figure; \texttt{blase} did not detect a spectral line here,
    but we have to artificially populate those points with lines that 
    have $\ln(a) = -1000$.}
\end{figure*}

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators generate multiple 
manifolds with the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 13.2 GB of disk space. Since this is now evaluable,
we call this the PHOENIX generator.

@@

\subsection{Performance Results}
Text

>>>>

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into an interpretable format.
We wanted to represent the spectra as a list of spectral lines rather than a
list of fluxes. This was done using \texttt{blase}, which detects spectral 
lines as Voigt profiles and tunes the profiles to mimic the original 
PHOENIX spectrum with back propagation. Four parameters were optimized: the 
line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
learning rate of 0.05 over 100 epochs. In addition, we limited two custom 
parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
that determines the extent of the Voigt profile to evaluate, saving
computational resources by not evaluating small numbers. Prominence sets a
lower limit for the amplitude of detected lines, which also saves resources 
by disregarding shallow lines. In short, larger values for wing cut and 
smaller values for prominence both increase the accuracy of using
\texttt{blase}'s line detection and optimization at the expense of an
increased computational cost. In addition, it should be mentioned that
\texttt{blase} uses the pseudo-Voigt approximation, which saves on 
computational cost while remaining accurate to about 1\%. It is a 
weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voight profile implementation uses the following:
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = \eta L(\lambda - \mu'; f) + (1 - \eta)G(\lambda - \mu'; f)\\
    \nonumber \eta = \sum_{n=1}^3 a_n\left(\dfrac{2\gamma}{f}\right)^n \qquad a = \mqty[1.36603 \\ -0.47719 \\ 0.11116]\\
    \nonumber f = 32\sum_{n=0}^5 b_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n \qquad b = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
\end{gather}
Notice that we use $\mu'$ instead of $\mu$ in the formula. This is because
\texttt{blase} allows the line center to shift slightly during optimization, 
and it is this shifted center which is used in computation. The individual 
Voigt profiles are still indexed by $\mu$ for cross-model line identification,
explained in the next section. Once optimization was complete, the list of 
identified lines, present in the state dictionary of \texttt{blase}'s model, 
was saved to a \texttt{.pt} file for each PHOENIX subset grid point. The total 
disk space these files took up is 465 MB.

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
Earlier we mentioned that \texttt{blase} tuned the line centers of
detected lines. This means that from one PHOENIX spectrum to the next, the
same line could have a slightly different line center. Since the goal of this
study is to interpolate the properties of each line, we needed to identify the
presence of a particular line across the PHOENIX subset. We decided to do
this by using the line centers $\mu$ of the detected lines pre-optimization. 
Now with each spectral line indexed by $\mu$, we had four parameters to interpolate:
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Note that since we are dealing with 
the parameters of Voigt profiles, we can see in Equation 2 that even if the 
interpolation method is linear, the actual result of interpolation varies 
non-linearly.

@@

\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{images/figure2.png}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX subset
    slice at solar metallicity. Notice the missing chunk in the top left
    of the figure; \texttt{blase} did not detect a spectral line here,
    but we have to artificially populate those points with lines that 
    have $\ln(a) = -1000$. The change in spacing on the x-axis is owing to
    the step of $T_\mathrm{eff}$ in the PHOENIX grid changing from 100 K to 
    200 K after $T_\mathrm{eff} = 7000$ K.}
\end{figure*}

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators generate multiple 
manifolds with the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 13.2 GB of disk space. This evaluation 
is able to reconstruct an existing PHOENIX spectrum or alternatively 
interpolate a new spectrum within the domain of the PHOENIX subset,
we call this the PHOENIX generator. In Figure 3, we show the same 
spectral line as in Figure 2, but now supersampled using the PHOENIX
generator evaluated over the same slice.

\begin{figure*}[t!]
    \centering
    \includegraphics[width=\textwidth]{images/figure3.png}
    \caption{Heatmap showing how $\ln(a)$ and $\sigma$ vary over the PHOENIX 
    subset slice at solar metallicity, now supersampled with the PHOENIX 
    generator. Notice that the missing chunk in the top left still exists 
    and does not display any artifacts, as the artificially populated points 
    are removed after interpolation to retain the model's integrity. Also
    see that the x-axis spacing is now uniform, as the PHOENIX generator 
    was evaluated at constant step.}
\end{figure*}

@@

\subsection{Spectral Reconstruction Time}
The typical use case for the PHOENIX generator may be to batch 
reconstruct spectra from an array of input coordinates. Therefore, there
is some motivation to attempt to reduce the computational cost of this
procedure to tractable levels. We evaluate the computational time needed
to use the PHOENIX generator in two distinct ways. First, for a single
input, which would be relevant in serial applications. Second, for an array
of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of 
an entire array of input coordinates to be evaluated at once, however 
using \texttt{blase} to reconstruct the spectrum from our interpolated
state dictionary is done serially, leading to what is actually more of 
a psuedo-parallel evaluation.

\subsection{Bayesian Optimizer Accuracy}
TACC
